---
title: "AutoRel Analysis Report"
author: "Generated by AutoRel Tool"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
params:
  output_path: "output/autorrel_results"
  contrast_group: "Group"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(pheatmap)
library(DT)

# Load results
results_file <- normalizePath(file.path(params$output_path, "all_results.csv"))
relevant_file <- normalizePath(file.path(params$output_path, "relevant_genes.csv"))

if (!file.exists(results_file)) {
  stop(paste("Results file not found at:", results_file))
}

results <- read.csv(results_file)
relevant <- read.csv(relevant_file)

# To create a volcano-style plot, we need to merge with original results (p-values)
# For the template, we'll look for results_dds_deseq2.rds in temp data
orig_res_path <- normalizePath(file.path(params$output_path, "temporary_data", "results_dds_deseq2.rds"), mustWork = FALSE)
if (file.exists(orig_res_path)) {
  orig_res <- readRDS(orig_res_path) %>% as.data.frame()
  plot_df <- results %>% inner_join(tibble::rownames_to_column(orig_res, "rowname"), by = "rowname")
} else {
  plot_df <- NULL
}
```

## Summary Findings

Total genes analyzed: **`r nrow(results)`**  
Total genes prioritized as **Relevant**: **`r nrow(relevant)`**

### Prioritization Table (Top 20)
```{r}
relevant %>% 
  head(20) %>% 
  dplyr::select(dplyr::any_of(c("rowname", "pred"))) %>% 
  datatable()
```

## Visualizations

### Relevance vs. Significance
This plot compares the AutoRel classification against the traditional statistical significance (p-values).

```{r, fig.width=8, fig.height=6}
if (!is.null(plot_df)) {
  ggplot(plot_df, aes(x = log2FoldChange, y = -log10(padj), color = pred)) +
    geom_point(alpha = 0.6) +
    scale_color_manual(values = c("relevant" = "#e41a1c", "irrelevant" = "#cccccc")) +
    theme_minimal() +
    labs(title = "AutoRel Classification Over Volcano Plot",
         subtitle = "Red points are prioritized as biologically relevant",
         x = "Log2 Fold Change",
         y = "-Log10 Adjusted P-value")
} else {
  print("Original DESeq2 results not found for visualization.")
}
```

### Expression Heatmap (Top Relevant)
Showing normalized expression levels for the most relevant genes.

```{r, fig.width=10, fig.height=8}
norm_counts_path <- file.path(params$output_path, "temporary_data", "norm_counts.rds")
if (file.exists(norm_counts_path)) {
  counts <- readRDS(norm_counts_path)
  top_genes <- relevant$rowname[1:min(20, nrow(relevant))]
  
  # Clean counts data frame (handle row names if needed)
  if (is.character(counts[,1])) {
    counts <- counts %>% column_to_rownames(colnames(counts)[1])
  }
  
  heatmap_mat <- counts[top_genes, ]
  
  pheatmap(log2(heatmap_mat + 1), 
           main = "Top 20 Relevant Genes (Log2 Normalized Counts)",
           color = colorRampPalette(c("blue", "white", "red"))(100),
           scale = "row",
           show_colnames = TRUE)
} else {
  print("Normalized counts not found for heatmap.")
}
```

---
*AutoRel Tool - Biological Relevance Prioritization*
